Linting Linearity in Core/System FC

### 2023-01-24 report.tex
Core Haskell

Haskel's intermediate representation that is statically typed

>>>>

Haskell is a big and expressive language with a lot of syntatic constructs and
many, many features. However, the whole of Haskell can be desugared down to a
very simple and elegant, explicitly typed, intermediate representation called
\textbf{Core}. To illustrate the difference in complexity, in GHC, the source
language (the big and complex Haskell) is defined through dozens of datatypes
and hundreds of constructors, while Core is defined in 3 types (expressions,
types, and coercions) and 15 constructors~\cite{}.
%
Core is unique as a strongly-typed intermediate language, with GHC as the only
production-ready compiler having such a typed intermediate representation.

Core allows us to reason about the entirety of Haskell in a much smaller and
elegant language -- analysis, optimizating transformations, and code generation
are all done on Core, not Haskell.
%
Of equal importance is Core being explicitly typed, which makes Core an internal
consistency tool that validates desugaring and optimization passes. Desugaring
and optimizing transformations must produce well-typed Core: having the Core
typechecker guarantee that Core programs are well-typed provides a verification
layer of the correctness of the optimizing transformations, of the desugarer,
and of their implementations.
%
Finally, the expressive but simple type-system in Core serves as a sanity-check
for all the extensions to the source language type system -- if we can desugar a
type system feature into Core then the feature must be sound by reducibility.

Core is typed very differently from Haskell because Core is explicitly typed and
its type system is based on the $System_{FC}$~\cite{systemfc} type system, while
Haskell is implicitly typed and its type system is based on the constraint-based
type inference system $OutsideIn(X)$~\cite{outsideinx}. Therefore, typechecking
Core is simple and fast and no inference must be done and, on the contrary,
Haskell's typechecker is much more complicated and requires ad-hoc restrictions
to make type inference feasible~\cite{}.
%
Haskell is typechecked in addition to Core to do type inference and to provide
meaningful type errors. If Haskell wasn't typechecked as it is, everything (e.g.
all binders) would have to be explicitly typed and all error messages would
refer to the intermediate language rather than the written program.

\todo[inline]{Any feature added to Haskell is only syntatic sugar}
